Influence of Connectivity on Dung Beetle Communities

### 2025-06-20 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
formatting

---
title: "Influence of Connectivity on Dung Beetle Communities"
author: "ERIC ESCOBAR-CHENA"
# date: "\\vspace{-1em}"
header-includes:
 # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
 - \usepackage{lineno}
 - \linenumbers
 - \usepackage[default]{sourcesanspro}
 - \usepackage{fancyhdr} # modify header into left, center, right
 - \pagestyle{fancy} # cues the use of fancy headers
 - \fancyfoot{} # cues the use of fancy footer
 - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
 - \fancyhead[R]{p. \thepage} # name and page on header right
 - \usepackage{subcaption}
 - \usepackage{float}
# # - \fancyfoot[R]{ }
 - \usepackage{setspace}
 # - doublespacing # manipulate line spacing 
# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
# - \usepackage{xcolor} # changes colors of text
# # http://latexcolor.com/ to choose colors
# # define the colors used above with rgb codes
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# # to center to top level (#) headings (my name) and change color of font
# # more info # https://www.ctan.org/pkg/sectsty
# - \usepackage{sectsty}
# - \sectionfont{\centering\color{darkmidnightblue}}
# # to center to sub-sub-headings (###) and change color of font
# - \usepackage{sectsty} 
# - \subsectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} 
# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
# - \usepackage{wrapfig} 
# - \newcommand{\blandscape}{\begin{landscape}}
# - \newcommand{\elandscape}{\end{landscape}}
# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
# - \usepackage{indentfirst}
# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
bibliography: ["eric-thesis.bib"]
output: bookdown::pdf_document2
  # pdf_document:
  #   includes:
  #     # in_header: style/preamble.tex
  #     template: null
  #     keep_tex: no
  #     extra_dependencies: ["float"]
      # number_sections: true
# documentclass: style/nsf2
biblio-style: apalike
csl: ecology.csl
---

@@

5) **Figure: ** biomass of different (R/T/D) functional groups.

@@

ERIC ESCOBAR-CHENA

A THESIS PRESENTED TO THE GRADUATE SCHOOL OF THE UNIVERSITY OF FLORIDA IN PARTIAL FULFILLMENT OF THE REQUIREMENTS FOR THE DEGREE OF MASTER OF SCIENCE

UNIVERSITY OF FLORIDA

2025

-----

© 2025 Eric Escobar-Chena

-----

<!-- _Your dedication is typed here. It should begin with the word “To.” (To my Mom is a typical dedication) If your dedication is longer than a single line it should be single-spaced, and centered vertically and horizontally_ -->

To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.

----

**ACKNOWLEDGMENTS**
<!-- _Acknowledgments must be written in complete sentences. Do not use direct address. For example, instead of Thanks, Mom and Dad!, you should say I thank my parents._  -->

@@

----

**LIST OF ABBREVIATIONS**

@@

----

Abstract of Thesis Presented to the Graduate School  
of the University of Florida in Partial Fulfillment of the  
Requirements for the Degree of Master of Science.

@@

Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation [@haddad_corridor_2003]. By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy  [@forman_general_1995].

[**EB edit:** break this next sentence into 3 to carry reader though the logic: changes in species diverstiy are important to understand because species are involved in interactions. If we gain or lose species we could gain or lose interactions and resulting ecosystems processes. there is some evidence of this, but... look at the green text notes below]

Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].

<!-- stuff that these inviduals or species do: -->
<!-- 1) their interactions with other species -->
<!-- - predation -->
<!-- - pollination -->
<!-- - seed dispersal -->
<!-- 2) stuff on their own that influences ecosystems -->
<!-- - ecosystem engineering (beavers, leaf-cutter ants) -->
<!-- - nutrient cycling -->
<!-- - decomposition -->

@@

Biodiversity between patch types was compared using Hill numbers, a set of indexes developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity [@jost_entropy_2006]. They are an alternative to more specialized metrics such as alpha, beta, and gamma diversities while being more standardized than other indexes such as Renyi or HCDT entropies, of which both groups of metrics are less intuitive for interpretation. Hill numbers are now the preferred metric for describing community dynamics for two reasons.  First, they are extrapolated from the same equation, manipulating a single parameter (i.e., _q_) to arrive at estimates of richness and diversity. Second, by  manipulating _q_ we can gain an understanding of compositional shifts otherwise obscured while using species richness [@chao_rarefaction_2014]. We compared community composition by increasing magnitudes of diversity components (i.e., _qD_) of _0D_ (i.e., species richness), _1D_ (i.e., Shannon entropy), and _2D_ (i.e, Simpson Diversity). <!--[can you add a sentence about how these are calculated? by which I mean - are the calculations different than for richness, shannon, etc.?]--> Diversity numbers and species richness were calculated using the package `hill` [@li_hillr_2018] for the R statistical programming language  [@posit_team_rstudio_2025]. Diversity numbers were calculated using package `iNEXT` [@hsieh_inext_2016]. Bray-Curtis dissimilarity values were calculated using package `Vegan` [@oksanen_vegan_2025]. Dung beetles were assigned traits by waste removal guild and habitat preference.

To test for the effects of connectivity on  abundance, species richness, and species diversity we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix. For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution [@bolker_generalized_2009]. Compared (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type. We included the identity of the sampling block as a random effects. To model our diversity metrics we took a similar approach, but this time using GLMMs with a Gaussian distribution [@chao_rarefaction_2014]. All models were fit using `lme4` package  [@bates_fitting_2015].  Prior to conducting our modeling we evaluated the the suitability of our data with _qqplots_ generated with the `DARMa` package  [@hartig_dharma_2024].

@@

<!-- 1. 1nd pass: roller, tunneler, dweller -->
<!-- 1. 2nd: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025) -->
<!-- 1. table: results of analyses  -->
<!-- 1. visualization: biomass of different (R/T/D) functional groups?    -->

@@

Overall, I collected N = `r total_N` dung beetles belonging to N = `r total_spp` species. The N = `r how_many_common` most dominant species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1). All but four species were captured in every patch type. _Onthophagus concinnus_ was only found in the matrix and winged patches, while _Onthophagus striatulus_ was only captured in matrix habitat and rectangular patches. _Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively). All species were within their native ranges.

@@

When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches($\beta$ = -0.402, P < 0.001, Table 4).

Statistical analysis of abundance focused on the six most abundant species. A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 7). The baseline abundance (intercept) corresponds to the abundance of _Aphodius alloblackburneus_ in connected patches. Compared to this baseline, _Canthon vigilans_, _Ateuchus lecontei_, and _Phanaeus igneus_ showed significantly higher abundance  ($\beta$ = 1.05, 0.73, 0.66 respectively; all p < 0.001), while _Onthophagus pennsylvanicus_ had lower abundance ($\beta$ = –0.80, p < 0.001).

Species-patch interactions indicated indicated species specific responses to treatments. In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix $\beta$ = –0.59, p < 0.001). _Ateuchus lecontei_ and _Phanaeus igneus_ showed increased effect on abundance in rectangle patches relative to expectations ($\beta$ = 0.40 and 0.47, respectively; p < 0.05). Random effects indicated moderate variability among blocks (standard deviation of intercept = 0.51).

@@

For Shannon Diversity, the estimated mean in Connected patches was 5.29 (SE = 0.66, t = 7.95). None of the alternative patch types showed statistically significant differences compared to Connected: Matrix ($\beta$ = –0.085, SE = 0.50, t = –0.17), Rectangle ($\beta$ = 0.15, SE = 0.50, t = 0.30), or Winged ($\beta$ = –0.34, SE = 0.50, t = –0.68).

Similarly, for Simpson's Diversity, the average value in Connected patches was 4.07 (SE = 0.64, t = 6.37). Again, none of the other patch types had significant effects: Matrix ($\beta$ = 0.11, SE = 0.41, t = 0.26), Rectangle ($\beta$ = 0.17, SE = 0.41, t = 0.40), and Winged ($\beta$ = –0.37, SE = 0.41, t = –0.91).

@@

_Don't repeat results...tell us:_

@@

We also observed that patch effects were not equally proportional for all species. The abundances of all dominant species were positively influenced by matrix habitat, but _Aphoidius alloblackburneus_ responded more positively than any other species. _Phanaeus igneus_ and _Ateuchus lecontei_ also had higher positive effects in rectangle patches.

@@

_here write about low sample sizes and reasons for why dung beetle communities would be so similar across treatment types. Dispersal, scale (roslin), maybe again a good reason to expect matrix as a source pop._

@@

> - Ospina 2018: wing shape py habitat preference and differences in wing shape determined by species groupings. Beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies). Large body beetles tending towards lower energy flight strategies. maybe justification for corridor preference?
> - stanbrook and king 2022: tunnelers preferring open habitat and tunnelers also contributing more towards dung removal. 
> - Conover 2019: dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types 
> - Gimenez Gomez et al 2021 -> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used. However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.

@@

 1. Abundances in matrix vs connected patch and why this could be happening
 - source pop to habiitat edge

@@

4. bray curtis hinting at similar land uses between corridor and winged patch
 - like julians paper corridors benefit certain populations and more fit populations are able to make better use

---

@@

1. Abundance 
 - reason for highest abundances in matrix and connected patch
 - lower abundance in rectangle hinting at fragmentation effects

2. Species Richness & Diversity
 - the role of hbitat connectivity in shaping community structure or lack thereof
 - why do we think connnected had lowest species richness??

3. Functional traits and ecological impacts
 - did corridoors favor a functional trait
 - why might there be a trait response 
 - implications for ecosystem processes
 - like duung removal papers seed dispersal and yep

4. Comparing to previous studies
- how are things aligning
- think about the biology

5. Limitations and future work
- potential confounding factors (seasonality, distance from edge, sampling methodology and temmporal variation)
- other directions to go (dispersal -> radar, secifically measuring changes in ecosystem services)

6. Takeaways for conservation and management
- dung beetles are robust
- what do think about corridor design and considerations for fragmented lanndscapes
- practical applications think about the beetles

@@

\smallskip  
::: {#refs}
:::

\newpage
## FIGURES & TABLES

<!-- (text - Nathan Fig of plots) -->
\newpage

@@

<!-- a<-readPNG("images/pitfalls.png") -->
<!-- b<-readPNG("images/srs-map.png") -->
<!-- grid.arrange(rasterGrob(a),rasterGrob(b),ncol=1) -->

@@

<!-- ```{r Figure_hill, echo=FALSE,message = FALSE,warning=FALSE} -->
<!-- library(iNEXT) -->
<!-- # -->
<!-- # btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv")) -->
<!-- # spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv")) -->

<!-- btl_sums_patch<-btl_data %>% -->
<!--   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% -->
<!--   group_by(species,patch,block) %>% -->
<!--   mutate(patch=case_when( -->
<!--     patch == "c" ~ "Corridor", -->
<!--     patch == "m" ~ "Matrix", -->
<!--     patch == "w" ~ "Winged", -->
<!--     patch == "r" ~ "Rectangle", -->
<!--     .default = as.character(patch))) %>% -->
<!--   summarize(n=sum(n, na.rm=TRUE)) %>% -->
<!--   arrange(patch,desc(n)) %>% -->
<!--   rename(sp_code=species, -->
<!--          n_patch=n) %>% -->
<!--   mutate(pb=paste(patch,block,sep="-")) %>%  -->
<!--   ungroup() %>%  -->
<!--   select(-patch,-block) %>%  -->
<!--   pivot_wider( -->
<!--     names_from = pb, -->
<!--     values_from = c(n_patch))  -->

<!-- div.data <- btl_sums_patch |> -->
<!--   column_to_rownames(var = "sp_code") -->

@@

<!-- z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance") -->
<!-- ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") + -->
<!--   theme_bw(base_size=18) -->

@@

<!-- ```{r sp_richness_patches, echo = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width = '100%'} -->
<!-- knitr::include_graphics("./corridor_beetles/corridor_docs/eric_ms_thesis/images/sp_richness_patches.png") -->
<!-- ``` -->

\newpage
```{r summary-table, echo=FALSE,message = FALSE,warning=FALSE}
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html

>>>>

---
title: "Influence of Connectivity on Dung Beetle Communities"
author: "ERIC ESCOBAR-CHENA"
# date: "\\vspace{-1em}"
# header-includes:
 # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
#  - \usepackage{lineno}
#  - \linenumbers
#  - \usepackage[default]{sourcesanspro}
#  - \usepackage{fancyhdr} # modify header into left, center, right
#  - \pagestyle{fancy} # cues the use of fancy headers
#  - \fancyfoot{} # cues the use of fancy footer
#  - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
#  - \fancyhead[R]{p. \thepage} # name and page on header right
#  - \usepackage{subcaption}
#  - \usepackage{float}
# # # - \fancyfoot[R]{ }
#  - \usepackage{setspace}

header-includes:
  - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
  - \usepackage{float} #to keep figure in the location placed instead of end
  - \floatplacement{figure}{H}
  - \usepackage{booktabs}
  - \raggedbottom
  - \usepackage{fancyhdr}
  - \pagestyle{plain}
  - \fancyfoot[C]{\thepage}
  - \fancyhead[L]{}
  - \fancyhead[R]{\small{EEC MS Thesis Draft}}
  - \thispagestyle{fancy}
  - \usepackage{tabu}
  - \usepackage{makecell}
  - \usepackage{pdflscape}
  - \usepackage{longtable}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \setlength\parindent{22pt}
  - \usepackage{indentfirst}
  - \renewenvironment{abstract}{}{} #suppress the abstract
  - \usepackage{setspace}

 # - doublespacing # manipulate line spacing 
# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
# - \usepackage{xcolor} # changes colors of text
# # http://latexcolor.com/ to choose colors
# # define the colors used above with rgb codes
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# # to center to top level (#) headings (my name) and change color of font
# # more info # https://www.ctan.org/pkg/sectsty
# - \usepackage{sectsty}
# - \sectionfont{\centering\color{darkmidnightblue}}
# # to center to sub-sub-headings (###) and change color of font
# - \usepackage{sectsty} 
# - \subsectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} 
# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
# - \usepackage{wrapfig} 
# - \newcommand{\blandscape}{\begin{landscape}}
# - \newcommand{\elandscape}{\end{landscape}}
# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
# - \usepackage{indentfirst}
# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
bibliography: ["eric-thesis.bib"]
output: bookdown::pdf_document2
  # pdf_document:
  #   includes:
  #     # in_header: style/preamble.tex
  #     template: null
  #     keep_tex: no
  #     extra_dependencies: ["float"]
      # number_sections: true
# documentclass: style/nsf2
biblio-style: apalike
csl: ecology.csl
---

@@

5)  **Figure:** biomass of different (R/T/D) functional groups.

@@

\vspace{8cm}

\begin{center}
THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES
\end{center}

\vspace{6cm}

\begin{center}
By
\end{center}

\vspace{3cm}

\begin{center}
ERIC ESCOBAR-CHENA
\end{center}

\vspace{5cm}

\begin{center}
A THESIS PRESENTED TO THE GRADUATE SCHOOL OF THE UNIVERSITY OF FLORIDA IN PARTIAL FULFILLMENT OF THE REQUIREMENTS FOR THE DEGREE OF MASTER OF SCIENCE
\end{center}

\vspace{2cm}

\begin{center}
2025
\end{center}

\vspace{20cm}

\begin{center}
© 2025 Eric Escobar-Chena
\end{center}

\vspace{15cm} 
<!-- _Your dedication is typed here. It should begin with the word “To.” (To my Mom is a typical dedication) If your dedication is longer than a single line it should be single-spaced, and centered vertically and horizontally_ -->

\begin{center}
To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.
\end{center}

\begin{center}

ACKNOWLEDGMENTS

\end{center}
<!-- _Acknowledgments must be written in complete sentences. Do not use direct address. For example, instead of Thanks, Mom and Dad!, you should say I thank my parents._  -->
\begin{center}

@@

\end{center}

\begin{center}

LIST OF ABBREVIATIONS

\end{center}

@@

\begin{center}

Abstract of Thesis Presented to the Graduate School

of the University of Florida in Partial Fulfillment of the

Requirements for the Degree of Master of Science.

\end{center}

@@

Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation [@haddad_corridor_2003]. By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995]. *[**EB edit:** break this next sentence into 3 to carry reader though the logic: changes in species diverstiy are important to understand because species are involved in interactions. If we gain or lose species we could gain or lose interactions and resulting ecosystems processes. there is some evidence of this, but... look at the green text notes below]* Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].

<!-- stuff that these inviduals or species do: -->

<!-- 1) their interactions with other species -->

<!-- - predation -->

<!-- - pollination -->

<!-- - seed dispersal -->

<!-- 2) stuff on their own that influences ecosystems -->

<!-- - ecosystem engineering (beavers, leaf-cutter ants) -->

<!-- - nutrient cycling -->

<!-- - decomposition -->

@@

Biodiversity between patch types was compared using Hill numbers, a set of indexes developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity [@jost_entropy_2006]. They are an alternative to more specialized metrics such as alpha, beta, and gamma diversities while being more standardized than other indexes such as Renyi or HCDT entropies, of which both groups of metrics are less intuitive for interpretation. Hill numbers are now the preferred metric for describing community dynamics for two reasons. First, they are extrapolated from the same equation, manipulating a single parameter (i.e., *q*) to arrive at estimates of richness and diversity. Second, by manipulating *q* we can gain an understanding of compositional shifts otherwise obscured while using species richness [@chao_rarefaction_2014]. We compared community composition by increasing magnitudes of diversity components (i.e., *qD*) of *0D* (i.e., species richness), *1D* (i.e., Shannon entropy), and *2D* (i.e, Simpson Diversity). <!--[can you add a sentence about how these are calculated? by which I mean - are the calculations different than for richness, shannon, etc.?]--> Diversity numbers and species richness were calculated using the package `hill` [@li_hillr_2018] for the R statistical programming language [@posit_team_rstudio_2025]. Diversity numbers were calculated using package `iNEXT` [@hsieh_inext_2016]. Bray-Curtis dissimilarity values were calculated using package `Vegan` [@oksanen_vegan_2025]. Dung beetles were assigned traits by waste removal guild and habitat preference.

To test for the effects of connectivity on abundance, species richness, and species diversity we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix. For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution [@bolker_generalized_2009]. Compared (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type. We included the identity of the sampling block as a random effects. To model our diversity metrics we took a similar approach, but this time using GLMMs with a Gaussian distribution [@chao_rarefaction_2014]. All models were fit using `lme4` package [@bates_fitting_2015]. Prior to conducting our modeling we evaluated the the suitability of our data with *qqplots* generated with the `DARMa` package [@hartig_dharma_2024].

@@

<!-- 1. 1nd pass: roller, tunneler, dweller -->

<!-- 1. 2nd: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025) -->

<!-- 1. table: results of analyses  -->

<!-- 1. visualization: biomass of different (R/T/D) functional groups?    -->

@@

Overall, I collected N = `r total_N` dung beetles belonging to N = `r total_spp` species. The N = `r how_many_common` most dominant species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1). All but four species were captured in every patch type. *Onthophagus concinnus* was only found in the matrix and winged patches, while *Onthophagus striatulus* was only captured in matrix habitat and rectangular patches. *Geotrupes blackburnii* and *Onthophagus tuberculifrons* were the only species restricted to one patch type (winged and matrix, respectively). All species were within their native ranges.

@@

When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches($\beta$ = -0.402, P \< 0.001, Table 4).

Statistical analysis of abundance focused on the six most abundant species. A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 7). The baseline abundance (intercept) corresponds to the abundance of *Aphodius alloblackburneus* in connected patches. Compared to this baseline, *Canthon vigilans*, *Ateuchus lecontei*, and *Phanaeus igneus* showed significantly higher abundance ($\beta$ = 1.05, 0.73, 0.66 respectively; all p \< 0.001), while *Onthophagus pennsylvanicus* had lower abundance ($\beta$ = –0.80, p \< 0.001).

Species-patch interactions indicated indicated species specific responses to treatments. In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix $\beta$ = –0.59, p \< 0.001). *Ateuchus lecontei* and *Phanaeus igneus* showed increased effect on abundance in rectangle patches relative to expectations ($\beta$ = 0.40 and 0.47, respectively; p \< 0.05). Random effects indicated moderate variability among blocks (standard deviation of intercept = 0.51).

@@

For Shannon Diversity, the estimated mean in Connected patches was 5.29 (SE = 0.66, t = 7.95). None of the alternative patch types showed statistically significant differences compared to Connected: Matrix ($\beta$ = –0.085, SE = 0.50, t = –0.17), Rectangle ($\beta$ = 0.15, SE = 0.50, t = 0.30), or Winged ($\beta$ = –0.34, SE = 0.50, t = –0.68). Similarly, for Simpson's Diversity, the average value in Connected patches was 4.07 (SE = 0.64, t = 6.37). Again, none of the other patch types had significant effects: Matrix ($\beta$ = 0.11, SE = 0.41, t = 0.26), Rectangle ($\beta$ = 0.17, SE = 0.41, t = 0.40), and Winged ($\beta$ = –0.37, SE = 0.41, t = –0.91).

@@

*Don't repeat results...tell us:*

@@

We also observed that patch effects were not equally proportional for all species. The abundances of all dominant species were positively influenced by matrix habitat, but *Aphoidius alloblackburneus* responded more positively than any other species. *Phanaeus igneus* and *Ateuchus lecontei* also had higher positive effects in rectangle patches.

@@

*here write about low sample sizes and reasons for why dung beetle communities would be so similar across treatment types. Dispersal, scale (roslin), maybe again a good reason to expect matrix as a source pop.*

@@

> -   Ospina 2018: wing shape py habitat preference and differences in wing shape determined by species groupings. Beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies). Large body beetles tending towards lower energy flight strategies. maybe justification for corridor preference?
> -   stanbrook and king 2022: tunnelers preferring open habitat and tunnelers also contributing more towards dung removal.
> -   Conover 2019: dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types
> -   Gimenez Gomez et al 2021 -\> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used. However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.

@@

1.  Abundances in matrix vs connected patch and why this could be happening

-   source pop to habiitat edge

@@

4.  bray curtis hinting at similar land uses between corridor and winged patch

-   like julians paper corridors benefit certain populations and more fit populations are able to make better use

------------------------------------------------------------------------

@@

1.  Abundance

-   reason for highest abundances in matrix and connected patch
-   lower abundance in rectangle hinting at fragmentation effects

2.  Species Richness & Diversity

-   the role of hbitat connectivity in shaping community structure or lack thereof
-   why do we think connnected had lowest species richness??

3.  Functional traits and ecological impacts

-   did corridoors favor a functional trait
-   why might there be a trait response
-   implications for ecosystem processes
-   like duung removal papers seed dispersal and yep

4.  Comparing to previous studies

-   how are things aligning
-   think about the biology

5.  Limitations and future work

-   potential confounding factors (seasonality, distance from edge, sampling methodology and temmporal variation)
-   other directions to go (dispersal -\> radar, secifically measuring changes in ecosystem services)

6.  Takeaways for conservation and management

-   dung beetles are robust
-   what do think about corridor design and considerations for fragmented lanndscapes
-   practical applications think about the beetles

@@

\smallskip

::: {#refs}
:::

## FIGURES & TABLES

<!-- (text - Nathan Fig of plots) -->

@@

<!-- a<-readPNG("images/pitfalls.png") -->

<!-- b<-readPNG("images/srs-map.png") -->

<!-- grid.arrange(rasterGrob(a),rasterGrob(b),ncol=1) -->

@@

<!-- ```{r Figure_hill, echo=FALSE,message = FALSE,warning=FALSE} -->

<!-- library(iNEXT) -->

<!-- # -->

<!-- # btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv")) -->

<!-- # spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv")) -->

<!-- btl_sums_patch<-btl_data %>% -->

<!--   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% -->

<!--   group_by(species,patch,block) %>% -->

<!--   mutate(patch=case_when( -->

<!--     patch == "c" ~ "Corridor", -->

<!--     patch == "m" ~ "Matrix", -->

<!--     patch == "w" ~ "Winged", -->

<!--     patch == "r" ~ "Rectangle", -->

<!--     .default = as.character(patch))) %>% -->

<!--   summarize(n=sum(n, na.rm=TRUE)) %>% -->

<!--   arrange(patch,desc(n)) %>% -->

<!--   rename(sp_code=species, -->

<!--          n_patch=n) %>% -->

<!--   mutate(pb=paste(patch,block,sep="-")) %>%  -->

<!--   ungroup() %>%  -->

<!--   select(-patch,-block) %>%  -->

<!--   pivot_wider( -->

<!--     names_from = pb, -->

<!--     values_from = c(n_patch))  -->

<!-- div.data <- btl_sums_patch |> -->

<!--   column_to_rownames(var = "sp_code") -->

@@

<!-- z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance") -->

<!-- ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") + -->

<!--   theme_bw(base_size=18) -->

@@

<!-- ```{r sp_richness_patches, echo = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width = '100%'} -->

<!-- knitr::include_graphics("./corridor_beetles/corridor_docs/eric_ms_thesis/images/sp_richness_patches.png") -->

```{r summary-table, echo=FALSE,message = FALSE,warning=FALSE}
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
